---
title: "Brownlows Bets"
author: "Karen Hernandez"
date: "September 9, 2020"
output: 
  html_document:
    number_sections: true
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
require(tidyverse)
require(knitr)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

 At the first day of class there are 35 students in Math Statistics class and professor Brownlow always proposes this bet to his students "I will wager $2 to your bet of $1 that at least two people in today's class have  the same birthday.  He tells me he has made that bet 637 times so far. Where do you think his profits are? Discuss this in a markdown document

# What are the fair odds 

Simulate the betting process 1000 times. Will characterize the results on the simulation.

### `r as.character(Sys.Date())`

```{r}
days=1:365
saveProfit=vector("numeric")
for(run in 1:1000) {
  
profit=0
for(i in 1:637){
  bdays=sample(days,size=35,replace=TRUE)
  n=length(unique(bdays))
  if(n<35) {
    profit=profit+1
  } else{profit=profit-2}
}
saveProfit[run]=profit
}

```


```{r}
hist(saveProfit)
print(mean(saveProfit))
```

It looks like Brownlow is making about $282.64 which is the mean.

```{r}
sortProfit=sort(saveProfit)
print(sortProfit[25])
print(sortProfit[975])
```

The 95% confidence interval shows that the value range from `r sortProfit[25]` to `r sortProfit[975]`. However if you run that code again, the values will continue to change.

We know now that 35 students is the wrong size for this 2:1 bet.


# What size of class makes the 2:1 bet fair ?

```{r}
days=1:365
saveProfit=vector("numeric")
for(run in 1:1000) {
  
profit=0
for(i in 1:637){
  bdays=sample(days,size=29,replace=TRUE)
  n=length(unique(bdays))
  if(n<29) {
    profit=profit+1
  } else{profit=profit-2}
}
saveProfit[run]=profit
}

```


```{r}
hist(saveProfit)
print(mean(saveProfit))
```

Based off this chart we can agree that 28 or 29 students is the right size for this 2:1 bet.

# In the Class and payoff as described what is a 95% confidence interval for Brownlows earnings ?

```{r}
days=1:365
saveProfit=vector("numeric")
for(run in 1:1000) {
  
profit=0
for(i in 1:637){
  bdays=sample(days,size=35,replace=TRUE)
  n=length(unique(bdays))
  if(n<35) {
    profit=profit+1
  } else{profit=profit-4.50}
}
saveProfit[run]=profit
}

```


```{r}
hist(saveProfit)
```


This chart shows us that in order for this to be a fair game with 35 students Dr. J must pay $4.50 each time he loses.



# Discussion and Conclusion

Class should be 28 or 29 students for payoff to be fair.

<!--Discussion and conclusion here.  If professor Brownlow is consistantly making or losing money how could he change his payout so that his profits center on zero?  Also discuss what size his class should be in order to have his current scheme balance at zero profit.!-->

